Copulas and their potential for ecology

### 2019-12-18 AER_Responses.tex
further progress
almost done addressing referee comments, except for possible need for a notation table

>>><***DAN to fill in>

@@

>>><DAN to do>

>>>>

>>>Thank you for this suggestion, which is a very good one. We did as requested and find 
the results sections seem more readable now to a less technicaly oriented audience.

@@

>>>We added the sentence "Whereas Anderson et al (2018) provide a specific methodological 
pathway, based partly on copulas, for the important problem of theanalysis of multivariate 
ecological population count data; we instead advocate more broadly for the application of 
copulas across ecology."

### 2019-12-18 Paper.Rmd
almost done addressing referee comments, except for possible need for a notation table

<!--bivariate example, model selection-->
For the soil C and N data (section \ref{Data}, Table \ref{tab_data_info}, 
Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
from `BiCopIndTest` in the `VineCopula` package). Non-independence is also visually apparent.
The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
and computed AIC and BIC weights and corresponding
lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
`r BivMS_res_Loecke$InfCritRes$copname[which.min(BivMS_res_Loecke$InfCritRes$AIC)]` was the 
minimum-AIC copula, with AIC `r round(min(BivMS_res_Loecke$InfCritRes$AIC),2)`, whereas the 
normal copula had a much higher AIC, `r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)`, answering the 
first part of Q1 for these data: these data have a dependence structure markedly distinct 
from that of a normal copula. Recall that AIC differences of $2$
or $3$ are considered meaningful and differences of $8$ or greater
are considered definitive, so that data provide essentially no support
for the higher-AIC model in that case [@burnham2003_modelselection].
Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
These values are distinct from what a normal copula would give, namely, 0, helping to answer 
the second part 
Q1 for these data. The values also differed substantially 
from each other, helping answer the third part of Q1. 
The numbers reflect stronger upper- than lower-tail
dependence, and this is also visible in the extreme upper-right
and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).

Our model selection procedures do not reveal whether tail dependence parameters are 
*significantly* different from $0$ and from each other. 
Furthermore, we caution that, while our model selection results do convincingly show 
non-normal copula structure, model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula 
(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
available from `BiCopGofTest`).
Our alternative nonparametric
approach, results detailed next, provides information about significance of tail
associations and asymmetries of tail associations.

The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
C and N data
were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
of the data,
separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
(Fig. \ref{fig_soilCN_nonparam}). Results showed that tail associations of data are stronger in 
both the lower and upper tails than would be expected under a null hypothesis of a normal copula (Fig. \ref{fig_soilCN_nonparam}B-F).
The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
on surrogates, separately in two 
comparisons (Table \ref{tab_soilCN_asym}). Results showed that
upper-tail association was significantly stronger than lower-tail association; i.e., C and N values are
more related when high than when low. These results
are consistent with the model selection results, but go beyond them by providing information
about statistical significance. Thus our results provide an
affirmative answer to Q1 for the soil C and N data; this answer is 
represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
in the right-most box in the middle row.

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes Q1 results 
for the bivariate datasets, with some details in the Appendices. The two 
variables were significantly related for all datasets 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, row 2). 
The normal copula was a poor fit compared to the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
difference between these fits was marginal.
Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
Often, either lower- or upper-tail dependence statistics differed substantially from 0 
(Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
helping to answer the second and third parts of Q1. Though again the model
selection results do not provide 
information on statistical significance of these differences, and are subject to the
caveat that, for some datasets, even the lowest-AIC copula was not an objectively good fit
(Table \ref{tab_bivar_summary}, rows 5 and 6), nonparametric results 
(Table \ref{tab_bivar_summary}, row 10) showed that tail associations
deviated significantly from what would be expected from a normal-copula
null model, and were also significantly asymmetric, except for the Cedar Creek data.
See Figs \ref{SM-fig_bmr_birds_nonparam}-\ref{SM-fig_cc2000_nonparam} for analogues to 
Fig. \ref{fig_soilCN_nonparam} for the bivariate datasets other than the soil C and N
data, see Tables \ref{SM-tab_bmr_birds_fit}-\ref{SM-tab_cc2000fit} for analogues to Table \ref{tab_soilCNfit},
and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse, though non-significantly. Our 
generally affirmative (except for Cedar Creek), 
empirically based answer to Q1 for the bivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names of 
the soil C and N and bird and mammal body mass versus metabolic rate
datasets, in the middle row of boxes.
We did the same analyses 
with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
2007, independence of biomass and Shannon's index could not be rejected; 
for 1997-1999, results were similar to 2000.

@@

<!--multivariate summary-->
Table \ref{tab_multivar_summary} summarizes Q1 results for
multivariate data. Results supported the 
conclusions that non-normal copula structure, non-normal tail dependence, and asymmetric tail 
associations were common, though not universal, answering 
Q1 in the affirmative for these data. 
Most location pairs were non-independent 
(Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
pairs had best-fitting copulas that were not the normal copula (Table \ref{tab_multivar_summary}, 
row 3), and AIC values of best-fitting copulas were, on average
across location pairs, between $2.714$ (*Ceratium furca* abundance data)
and $5.764$ (leaf-curling plum aphid first-flight data) lower than
AIC values for the normal copula (Table \ref{tab_multivar_summary}, 
row 4). Best-fitting copulas were nearly always considered an adequate fit 
(Table \ref{tab_multivar_summary}, row 5). Some of the datasets (green spruce aphid
abundance, \textit{Ceratium furca} abundance, methane-flux) showed stronger 
lower- than upper-tail dependence 
(Table \ref{tab_multivar_summary}, rows 6-8), whereas leaf-curling plum aphid first flight
data showed the reverse. All of the values in Table \ref{tab_multivar_summary}, rows 8 deviated highly significantly
from what would have been expected under a null hypothesis 
of equal chances for positive and negative values.

@@

Our model with asymmetric environmental effects 
produced outputs with visually apparent asymmetry 
of tail associations between the ecological variables $E_i$ in the two locations, 
to an extent that depended on the value of $b$.
For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}A,B), 
for both $g_1$ and $g_2$,
asymmetry of tail association was strong; for 
$b=\pm 0.5$, asymmetry was weaker but still apparent
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}C-F).
Lower-tail (respectively, upper-tail) spatial associations in the 
effects of noise ($g_1$, respectively, $g_2$) 
produced lower-tail (respectively, upper-tail) spatial 
associations in the ecological variable, $E_i$.
Results using our statistic $\cor_{0,0.1}-\cor_{0.9,1}$ 
strongly reflected the visually apparent asymmetry (Table \ref{tab_stat_npa_asym}). 
Thus asymmetry of environmental effects is a mechanism that may be partly 
responsible for non-normal copula structure and
asymmetric tail associations across space in ecological variables.
This result is represented in Fig. \ref{MasterFigure} as the solid box around 
"Nonlinear environmental effects, Liebig's law" and the arrows labelled "A".
It is explained in the Discussion why the box and some of the arrows 
are solid, instead of dashed, although the results of this section are theoretical.

@@

<!--Second mechanism, the Moran one-->
The hypothesis that characteristics of 
the copula structure of spatial dependence 
in an ecological variable can be inherited from an environmental 
variable was found to be reasonable, because it held for the models we examined; 
however, similarities between environmental- and 
ecological-variable copula structure were reduced when 
dynamics were strongly nonlinear.
For simulations using AR(1) models, our correlation and tail asymmetry statistics 
were always similar for noise and populations 
(Fig. \ref{fig_Cause4copula_C_kend_ar1} for the Clayton copula, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_ar1} for the survival Clayton copula).
Though there were significant differences for many statistics
and simulations, these were small compared to the overall tendency for
larger (respectively, smaller) values of our statistics, as applied to
noise, to be paired with larger (respectively, smaller) values of the same
statistics as applied to model outputs. 
For our nonlinear model with weak noise, values of the statistics
were again quite similar for noise and model outputs 
(Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1} for Clayton, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_0.1} for survival Clayton).
Though there were again significant differences for many statistics
and simulations, these were again small relative to overall variation of 
values of the statistics.
Since many ecological models are nonlinear, this result provides the 
reasonable expectation that there may be a Moran-effect-like correspondence 
between noise and model-output dependence structure across space 
for typical ecological dynamics,
as long as environmental noise is small enough that dynamics stay relatively close
to the model equilibrium. Theoretical results that hold for "weak noise" in this sense
are common in ecology.
For strong noise and using our nonlinear model, 
our correlation and tail-association-asymmetry statistics, generally 
speaking, were approximately similar between noise and model
outputs; however, similarity was reduced relative to previous results, 
and for a few simulations, asymmetry statistics
had opposite signs for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1} and \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_1}). For instance, 
using a Clayton  copula with a large Kendall $\tau$,
$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}).

@@

\noindent Empirical results were consistent with the hypothesis that the skewness
of a spatial-average time series is influenced by tail associations
between the local quantities being averaged. 
For datasets that
exhibited highly asymmetric tail associations in earlier analyses (green spruce 
aphid abundance data had stronger lower-tail association, and, 
respectively,  leaf-curling plum 
aphid first flight data had stronger upper-tail association, Table 
\ref{tab_multivar_summary}), skewness of the spatial average 
was less than (respectively, greater than) a significant fraction of 
surrogate skewness values 
(Fig. \ref{fig_skewness_spearman}A,B). For datasets with moderately stronger 
lower than upper-tail 
dependence (\emph{Ceratium furca} abundance and methane data), 
skewness of the spatial
average showed a non-significant or marginally significant tendency 
toward being less than surrogate skewnesses (Fig. \ref{fig_skewness_spearman}C,D). 
Thus copula 
structure and asymmetric tail associations are important for spatially averaged 
quantities. This result is represented in Fig. \ref{MasterFigure}
as the solid box around "Instability/skewness of mean or total 
time series" and the solid arrows labeled "X".

>>>>

<!--bivariate example, model selection-->
For the soil C and N data (section \ref{Data}, Table \ref{tab_data_info}, 
Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
from `BiCopIndTest` in the `VineCopula` package). Non-independence is also visually apparent.
The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
and computed AIC and BIC weights and corresponding
lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
`r BivMS_res_Loecke$InfCritRes$copname[which.min(BivMS_res_Loecke$InfCritRes$AIC)]` was the 
minimum-AIC copula, with AIC `r round(min(BivMS_res_Loecke$InfCritRes$AIC),2)`, whereas the 
normal copula had a much higher AIC, `r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)`, answering the 
first part of Q1 for these data: the data have a dependence structure markedly distinct 
from that of a normal copula. Recall that AIC differences of $2$
or $3$ are considered meaningful and differences of $8$ or greater
are considered definitive, so that data provide essentially no support
for the higher-AIC model in that case [@burnham2003_modelselection].
Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
These values are distinct from what a normal copula would give, namely, 0, helping to answer 
the second part of
Q1 for these data. The values also differed substantially 
from each other, helping answer the third part of Q1. 
The numbers reflect stronger upper- than lower-tail
dependence, and this is also visible in the extreme upper-right
and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).
Apparently C and N are more strongly associated with each other in soils
for which they are both plentiful than in soils for which they are 
both less abundant, and C and N are more strongly associated in both
tails than would be modelled using a normal copula model or methods that
assume such a model.

However, our model selection procedures do not reveal whether tail dependence parameters are 
*significantly* different from $0$ and from each other. 
Furthermore, we caution that, while our model selection results do convincingly show 
non-normal copula structure, model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula 
(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
available from `BiCopGofTest`).
Our alternative nonparametric
approach, results detailed next, provides information about significance of tail
associations and asymmetries of tail associations.

The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
C and N data
were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
of the data,
separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
(Fig. \ref{fig_soilCN_nonparam}). Results confirmed that tail associations of data are stronger in 
both the lower and upper tails than would be expected under a null hypothesis of a normal copula (Fig. \ref{fig_soilCN_nonparam}B-F).
The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
on surrogates, separately in two 
comparisons (Table \ref{tab_soilCN_asym}). Results confirmed that
upper-tail association was significantly stronger than lower-tail association; i.e., C and N values are
more related when high than when low. These results
are consistent with the model selection results, but go beyond them by providing information
about statistical significance. Thus our results provide an
affirmative answer to Q1 for the soil C and N data; this answer is 
represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
in the right-most box in the middle row.

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes Q1 results 
for the bivariate datasets, with some details in the Appendices. The two 
variables were significantly related for all datasets 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, row 2). 
The normal copula was a poor fit compared to the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
difference between these fits was marginal.
Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
Often, either lower- or upper-tail dependence statistics differed substantially from 0 
(Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
helping to answer the second and third parts of Q1. Though again the model
selection results do not provide 
information on statistical significance of these differences, and are subject to the
caveat that, for some datasets, even the lowest-AIC copula was not an objectively good fit
(Table \ref{tab_bivar_summary}, rows 5 and 6), nonparametric results 
(Table \ref{tab_bivar_summary}, row 10) showed that tail associations
deviated significantly from what would be expected from a normal-copula
null model, and were also significantly asymmetric, except for the Cedar Creek data.
See Figs \ref{SM-fig_bmr_birds_nonparam}-\ref{SM-fig_cc2000_nonparam} for analogues to 
Fig. \ref{fig_soilCN_nonparam} for the bivariate datasets other than the soil C and N
data, see Tables \ref{SM-tab_bmr_birds_fit}-\ref{SM-tab_cc2000fit} for analogues to Table \ref{tab_soilCNfit},
and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse, though non-significantly. 
Apparently species body masses and metabolic rates, for both birds and mammals, are more strongly 
associated with each other when both these quantities are large than when they are 
both small. Our generally affirmative (except for Cedar Creek), 
empirically based answer to Q1 for the bivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names of 
the soil C and N and bird and mammal body mass versus metabolic rate
datasets, in the middle row of boxes.
We did the same analyses 
with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
2007, independence of biomass and Shannon's index could not be rejected; 
for 1997-1999, results were similar to 2000.

@@

<!--multivariate summary-->
Table \ref{tab_multivar_summary} summarizes Q1 results for
multivariate data. Results supported the 
conclusions that non-normal copula structure, non-normal tail dependence, and asymmetric tail 
associations were common, though not universal, answering 
Q1 in the affirmative for these data. 
Most location pairs were non-independent 
(Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
pairs had best-fitting copulas that were not the normal copula (Table \ref{tab_multivar_summary}, 
row 3), and AIC values of best-fitting copulas were, on average
across location pairs, between $2.714$ (*Ceratium furca* abundance data)
and $5.764$ (leaf-curling plum aphid first-flight data) lower than
AIC values for the normal copula (Table \ref{tab_multivar_summary}, 
row 4). Best-fitting copulas were nearly always considered an adequate fit 
(Table \ref{tab_multivar_summary}, row 5). Some of the datasets (green spruce aphid
abundance, \textit{Ceratium furca} abundance, methane-flux) showed stronger 
lower- than upper-tail dependence 
(Table \ref{tab_multivar_summary}, rows 6-8), whereas leaf-curling plum aphid first flight
data showed the reverse. 
Thus the spatial synchrony of rarity was stronger than that of outbreaks for 
green spruce aphid and \textit{Ceratium furca} abundance, whereas leaf-curling aphid
first flights were more spatially related to each other when these flights happened
late than when they happened early.
All of the values in Table \ref{tab_multivar_summary}, rows 8 deviated highly significantly
from what would have been expected under a null hypothesis 
of equal chances for positive and negative values.

@@

Our model with asymmetric environmental effects 
produced outputs with visually apparent asymmetry 
of tail associations between the ecological variables $E_i$ in the two locations, 
to an extent that depended on the value of $b$.
For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}A,B), 
for both $g_1$ and $g_2$,
asymmetry of tail association was strong; for 
$b=\pm 0.5$, asymmetry was weaker but still apparent
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}C-F).
Lower-tail (respectively, upper-tail) spatial associations in the 
effects of noise ($g_1$, respectively, $g_2$) 
produced lower-tail (respectively, upper-tail) spatial 
associations in the ecological variable, $E_i$.
Results using our statistic $\cor_{0,0.1}-\cor_{0.9,1}$ 
strongly reflected the visually apparent asymmetry (Table \ref{tab_stat_npa_asym}). 
Thus asymmetry of environmental effects is a mechanism that may be partly 
responsible for non-normal copula structure and
asymmetric tail associations across space in ecological variables:
when populations or other ecological quantities are influenced principally by
low (respectively, high) values of an environmental variable, one may expect 
left-tail (respectively, right-tail) associations across space for the ecological variable. 
This result is represented in Fig. \ref{MasterFigure} as the solid box around 
"Nonlinear environmental effects, Liebig's law" and the arrows labelled "A".
It is explained in the Discussion why the box and some of the arrows 
are solid, instead of dashed, although the results of this section are theoretical.

@@

<!--Second mechanism, the Moran one-->
The hypothesis that characteristics of 
the copula structure of spatial dependence 
in an ecological variable can be inherited from an environmental 
variable was found to be reasonable, because it held for the models we examined -
the Moran effect seems to extend to copula structure and tail associations - 
however, similarities between environmental- and 
ecological-variable copula structure were reduced when 
dynamics were strongly nonlinear.
For simulations using AR(1) models, our correlation and tail asymmetry statistics 
were always similar for noise and populations 
(Fig. \ref{fig_Cause4copula_C_kend_ar1} for the Clayton copula, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_ar1} for the survival Clayton copula).
Though there were significant differences for many statistics
and simulations, these were small compared to the overall tendency for
larger (respectively, smaller) values of our statistics, as applied to
noise, to be paired with larger (respectively, smaller) values of the same
statistics as applied to model outputs. 
For our nonlinear model with weak noise, values of the statistics
were again quite similar for noise and model outputs 
(Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1} for Clayton, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_0.1} for survival Clayton).
Though there were again significant differences for many statistics
and simulations, these were again small relative to overall variation of 
values of the statistics.
Since many ecological models are nonlinear, this result provides the 
reasonable expectation for a Moran-effect-like correspondence 
between noise and model-output dependence structure across space 
for typical ecological dynamics,
as long as environmental noise is small enough that dynamics stay relatively close
to the model equilibrium. Theoretical results that hold for "weak noise" in this sense
are common in ecology.
For strong noise and using our nonlinear model, 
our correlation and tail-association-asymmetry statistics, generally 
speaking, were approximately similar between noise and model
outputs; however, similarity was reduced relative to previous results, 
and for a few simulations, asymmetry statistics
had opposite signs for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1} and \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_1}). For instance, 
using a Clayton  copula with a large Kendall $\tau$,
$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}).

@@

\noindent Empirical results were consistent with the hypothesis that the skewness
of a spatial-average time series is influenced by tail associations
between the local quantities being averaged. 
For datasets that
exhibited highly asymmetric tail associations in earlier analyses (green spruce 
aphid abundance data had stronger lower-tail association, and, 
respectively,  leaf-curling plum 
aphid first flight data had stronger upper-tail association, Table 
\ref{tab_multivar_summary}), skewness of the spatial average 
was less than (respectively, greater than) a significant fraction of 
surrogate skewness values 
(Fig. \ref{fig_skewness_spearman}A,B). For datasets with moderately stronger 
lower than upper-tail 
dependence (\emph{Ceratium furca} abundance and methane data), 
skewness of the spatial
average showed a non-significant or marginally significant tendency 
toward being less than surrogate skewnesses (Fig. \ref{fig_skewness_spearman}C,D). 
Thus copula 
structure and asymmetric tail associations are important for spatially averaged 
quantities and can influence the temporal stability of those quantities. 
This result is represented in Fig. \ref{MasterFigure}
as the solid box around "Instability/skewness of mean or total 
time series" and the solid arrows labeled "X".
